Manuscript drafts

### 2020-05-11 manuscript/drafts/manuscript.Rmd

---
title: "Manuscript drafts"
author: "Gerbrich Ferdinands"
date: "22/04/2020"
output:
  word_document: default
  citation_package: biblatex
  pdf_document:
bibliography: asreview.bib
csl: systematic-reviews.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

options(tinytex.verbose = TRUE)

library(tidyverse)
library(knitr)
library(xtable)
library(kableExtra)
library(captioner)
library(glue)
library(rbbt)

# load model parameters 
source("asr-parameters.R")

@@

# Introduction
Systematic reviews are top of the bill in building evidence in research. A systematic review brings together all studies relevant to answer a specific research question [@PRISMA-PGroup2015]. Systematic reviews inform practice and policy [@Gough2002] and are key in developing clinical guidelines [@Chalmers2007]. 
<!-- PRISMA definition:  -->
<!-- > "A systematic review attempts to collate all relevant evidences that fits pre-specified eligibility criteria to answer a specific research -->
<!-- question. 
add that this is a systematic/rigorous process?-->
However, systematic reviews are costly because they involve the manual screening of thousands of titles and abstracts, identifying publications relevant to answering the research question. An experienced reviewer takes on average 30 seconds to screen one title and abstract, whereas an inexperienced reviewer takes even longer [@Wallace2010]. Conducting a systematic review typically requires over a year of work by a team of researchers [@Borah2017]. 
<!-- Besides time-consuming, the process of title and abstract screening is prone to human error [@Wang2020]. -->
<!-- (14 to 77,910, trimmed mean 1,286).  -->
<!-- Need to be screened by multiple researchers.  -->

Moreover, systematic reviewers are often bound to a limited budget and timeframe. Currently, the demand for systematic reviews exceeds the available time and resources by far [@Lau2019]. Especially when the need for guidelines is urgent - such as in the context of the current COVID-19 crisis - it is almost impossible to provide a review that is both timely and comprehensive. To ensure a timely review, reducing workload in systematic reviews is imperative.

With advances in Artificial Intelligence (AI), there has been wide interest in tools to reduce workload in systematic reviews [@Harrison2020]. Various learning models have been proposed, aiming to predict whether a given publication is relevant or irrelevant to the systematic review. Findings suggest that such models potentially reduce workload with 30-70% at the cost of losing 5% of relevant publications (95% recall) [@OMara-Eves2015].
<!-- To date, several studies have begun to examine the use of learning algorithms for screening prioritization. -->

Screening prioritization is a well-established approach in increasing efficiency in title and abstract screening [@Cohen2009; @Shemilt2014]. In screening prioritization, the learning model reorders publications to be screened by their likeliness to be relevant. The model presents the reviewer with the publications which are most likely to be relevant first, thereby expediting the process of finding all of the relevant publications. Such an approach allows for substantial time-savings in the screening process. Reviewing relevant publications early facilitates a faster transition of those publications to the next steps in the review process [@Cohen2009]. Additionally, several studies report increasing efficiency beyond saving time [@OMara-Eves2015].

<!-- assist the reviewer in the screening process -->
<!-- Several studies have proposed active learning models for screening prioritization -->
Recent studies have demonstrated the effectiveness of screening prioritization by means of active learning models [@Yu2018; @Yu2018; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010]. Active learning is when the model can iteratively improve its predictions by allowing the model to choose the data from which it can learn [@Settles2012]. Active learning has proven to be an efficient strategy in large datasets where labels are scarce, which makes title and abstract screening an ideal candidate for such models. When applied in screening prioritization, the reviewer screens publications that are presented by an active learning model. Subsequently, the active learning model learns from the reviewers' decision ('relevant', 'irrelevant') and uses this knowledge in selecting the next publication to be screened by the reviewer.

Although prior studies on the application of active learning models in title and abstract screening have addressed important issues [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010], the complex nature of the field is making it difficult to draw overarching conclusions about best practice [@OMara-Eves2015]. First, whilst there exists a scala of classification techniques, research to date has not yet evaluated techniques other than Support Vector Machines (SVM) [@Yu2018; @Yu2018; @Miwa2014; @Wallace2010]. Second, Miwa et al. [@Miwa2014] were the only study so far to make a direct comparison of model performance across research areass and found that active learning was more difficult on data from the social sciences compared to data from the medical sciences. The lack of replication on data from varying research contexts makes it impossible to draw conclusions about the general effectiveness of such techniques [@OMara-Eves2015; @Marshall2020]. The question remains how active learning models for screening prioritization perform across different (1) classification techniques and (2) review contexts. Hence, additional evaluations of active learning models are required.

@@

The purpose of the current paper is to increase the evidence base of active learning models for reducing workload in title and abstract screening in systematic reviews. Combining latest insights from this area, we propose seven different active learning models for the purpose of identifying relevant publications in systematic review datasets. The models were chosen to maximize the number of identified relevant publications, while minimizing the number of publications needed to screen. 
Working towards a general consensus in this emerging field, model performance was asessed by conducting a retrospective simulation on six systematic review datasets. Datasets were collected from the fields of medicine, software engineering, psychology, behavioural public administration, and virology to assess generalizability of the models across research contexts. The models, datasets and simulations are implemented in a pipeline of active learning for prioritization screening, called $\texttt{ASReview}$ [@ASReview2020]. $\texttt{ASReview}$ is an open source and generic tool such that users can adapt and add modules as they like, encouraging fellow researchers to replicate findings from previous studies. All scripts and data used are openly published to facilitate usability and acceptability of AI-assisted title and abstract screening in the field of systematic review.

The remaining part of this paper is organized as follows. The Technical Details section will cover the workings of active learning models for study selection in systematic reviews on a conceptual level. 
The method section describes ... The results section reports ... The discusssion section summarises the findings, comments on them and  summarises the main findings, discusses 
discuss limitations, draw conclusion and

@@

# Methods 
## Task Description
The screening process of a systematic review starts with all publications obtained in the search. The task is to identify which of these publications are relevant, by screening the publications at the title and abstract level.  In active learning for screening prioritiztaion, the screening process proceeds as follows:

- Start with the set of all publications $x$ with an unknown label, $\mathcal{U}$.
- The reviewer provides a label for a few publications $x \in \mathcal{U}$, creating an set of labeled publications $\mathcal{L}$. The publications can be labeled as relevant $x_R$ or irrelevant $x_I$.
- The active learning cycle starts:

  1. A classifier is trained on the labeled publications, $C = \texttt{train}(\mathcal{L})$

  2. The classifier predicts labels for all unlabelled publications, $C(\mathcal{U})$

  3. Based on the predictions by $C$, the model selects the most relevant publication $x \in \mathcal{U}$

  4. The model asks the reviewer to screen this publication, $x_?$

  5. The reviewer screens the publication and provides a label, $x_R$ or $x_I$

  6. The labeled publication is added to the training data, $\mathcal{L}$

@@

- In this active learning cycle, the model can incrementally improve its predictions on the remaining unlabeled publications. The relevant publications are identified as early in the process as possible. The reviewer and the model keep interacting until the reviewer decides to stop or until all publications in $\mathbf{U}$ have been labelled.

@@

#### Classification
To make predictions on the unlabeled publications, a classifier is trained on features from the set of previously labeled publications. 
<!-- The classifier predicts class of a given unlabeled publication.  -->
A technique widely used in classification tasks is the Support Vector Machine (SVM). SVMs separate the data into classes by finding a multidimensional hyperplane [@Tong2001; @Kremer2014]. SVMs have been proven to be effective in active learning models for screening prioritization [@Yu2018; Miwa2014]. Moreover, SVMs are the currently the only classifier implemented in ready-to-use software tools implementing active learning for screening prioritization (Abstrackr [@Wallace2012b], Colandr [@Cheng2018], FASTREAD [@Yu2018], Rayyan [@Ouzzani2016], and RobotAnalyst [@Przybyla2018]).

Whilst performance several classification techniques have been investigated in the AI-aided title and abstract screening field in general [SOURCE], the relatively new subfield of active learning for screening prioritization has not yet studied the performance of classifiers other than SVMs [@Yu2018; @Yu2018; @Cormack2014; @Cormack2015; @Miwa2014; @Wallace2010]. 
<!-- Whilst the SVM is a well-known technique, also classifiers have been employed in class prediction tasks in systematic reviews.  -->
<!-- In fact, research in this area to date has not yet determined the performance of classifiers other than SVMs. -->
The current study aims to address this gap by exploring performance of three classifiers besides SVM:

@@

These three classification techniques were selected because they are all widely adopted in previous studies on text classifcation [source]. Moreover, the amount of processing power is relatively low. They can run on a personal computer.

@@

#### Class imbalance problem
There are two classes in the dataset: relevant and irrelevant publications. Typically, only a fraction of the publications in the data belong to the relevant class. This poses a problem for training a classifier as there are far fewer examples of relevant than irrelevant publications to train on [@OMara-Eves2015]. Moreover, classifiers are well-suited to separate data into classes, but not to correctly identifying one class [@Wallace2010]. This is evident in the case of a dataset where only one percent of publications are relevant. A model would achieve 99% accuracy when classifying all publications as irrelevant, even though none of the relevant papers would have been identified. Therefore, the class imbalance problem causes the classifier to miss relevant publications.

Previous studies have addressed the class imbalance problem by rebalancing the training data in different ways [@OMara-Eves2015]. To decrease the class imbalance in the training data, the models in the current study rebalance the training set by Dynamic Supersampling (DS). DS decreases the number of irrelevant publications in the training data, whereas the number of relevant publications are increased (by copy) such that the size of the training data remains the same. The ratio between relevant and irrelevant publications is not fixed, but dynamically updated and depends on the size of the training data, the total number of publications, and the ratio between total number of relevant and irrelevant publications.

#### Word embeddings 
To predict publication class, the classifier uses information from the publications in the dataset. 
Examples of such information are titles and abstracts. However, a model cannot predict class from the titles and abstracts as they are; their textual content needs to be repressented numerically. The textual information needs to be mapped to feature vectors. This process of numerically representing textual content is called 'word embeddings'.

A classical example of word embeddings is a 'bag of words' (bow) representation. For each text in the data set, the number of occurrences of each word is stored. This leads to $n$ features, where $n$ is the number of distinct words in the texts [@scikit-learn]. The bag-of-words method is simplistic and will highly value often occuring but otherwise meaningless words such as "and". A more sophisticated approach is Term-frequency Inverse Document Frequency (TF-IDF). TF-IDF circumvents this problem by adjusting the term frequency in a text with the inverse docuement frequency, the frequency of a given word in the entire data set [@Ramos2003]. A downside of TF-IDF and other bow methods is that they do not take into account the ordering of the words, thereby ignoring semantics. An example of an approach that aims to overcome this weakness is Doc2Vec (D2V), capable of grasping the relations between words by learning to predict the words in the texts [@Le2014].

Miwa et al. found that active learning was more difficult on data from the social sciences compared to data the medical sciences and were able to link this difficulty to a natural difference in text complexity between these research areas [@Miwa2014]. As the study by Miwa et al. adopted a bow approach [@Miwa2014], we hypothesize that a more sophisticated word embeddings strategy has the potential to bridge the performance gap between these research areas.

#### Query strategy
The active learning model can adopt different strategies in selecting the next publication to be screened by the reviewer. 
A strategy mentioned before is selecting the publication with the highest probability of being relevant. In the active learning literature this is referred to as certainty-based active learning [@Settles2012]. Another well-known strategy is uncertainty-based active learning, where the instances that will be presented next will be those instances on which on which the model's classifications are the least certain, i.e. close to 0.5 probability [@Settles2012]. Traditionally, this strategy trains the most accurate model because the model can learn the most from instances it is uncertain about.

Even though uncertainty-based active learning generally leads to a more accurate model in the end, certainty-based active learning is the preferred strategy for the task at hand. The main reason is that this strategy is far better suited to the goal of prioritizing the relevant publications. Moreover, a study comparing performance of both strategies in found that the accuracy gain of uncertainty-based screening was not significant [@Miwa2014]. Furthermore, uncertainty-based active learning is far better equipped at dealing with imbalanced data in active learning [@Fu2011]. 
Therefore, certainty-based sampling is most equipped to the current scenario in which we are dealing with highly imbalanced datasets and where the goal is to identify all relevant publications as soon as possible.

@@

First, four models combining every classifier with TF-IDF word embeddings were investigated: 
 - SVM + TF-IDF
 - NB + TF-IDF
 - RF + TF-IDF 
 - SVM + TF-IDF

Second, the classifiers were combined with Doc2Vec word embeddings, leading to the following three models:^[The combination NB + D2V is impossible because NB can only process positive input features, whereas D2V]

@@

## Simulation study
For each of the seven models, performance was evaluated by simulating the model on the screening process of six systematic reviews. 
Performance of the seven models was evaluated by simulating their behaviour in the screening process of six systematic review datasets.
Put differently, 42 simulations were carried out. For every model - dataset combination, hyperparameters were optimized^[see the Appendix for more information]. To account for variance, every simulation was repeated for 15 trials. 
Simulations were run using ASReview's simulation mode [@ASReview2020]. There was no need for a human reviewer as the model could query the labels in the data instead.

Every simulation started with an initial training set of one relevant and one irrelevant publication to represent a 'worst case scenario' where the reviewer has minimal prior knowledge on the publications in the data. To account for bias, the initial training set was randomly sampled from the dataset for every of the 15 runs. Although varying over runs, the initial training sets were kept constant over datasets to allow for a direct comparison of models within datasets. A seed value was set to ensure reproducability. 
The classifier was retrained every time after a publication had been labeled. 
The simulation ended after all publications in the dataset had been labeled.

@@

## Datasets 
The models were simulated on a convenience sample of six systematic review datasets. The data selection process was driven by two factors. Firstly, datasets were selected based on their background, given the need for datasets from diverse research areas. Secondly, datasets were selected by their availability, given the limited timespan of the current project. The datasets were retrieved from a collection of open systematic review datasets to be used for text mining purposes^[https://github.com/asreview/systematic-review-datasets, [@ASReview2020]].

@@

Data were preprocessed from their original source into a test dataset, containing title and abstract of the publications obtained in the initial search. Candidate studies with missing abstracts and duplicate instances were removed from the data. Test datasets were labelled to indicate which candidate studies were included in the systematic review, thereby indicating relevant publications. All test datasets consisted of thousands of candidate studies, of which only only a fraction was deemed relevant to the systematic review. Inclusion rates were centered around 1-2 percent with one outlier of about 5 percent (`r table_nums("datasets", display = "cite")`). 
<!-- The data preprocessing scripts can be found on the GitHub^[https://github.com/GerbrichFerdinands/asreview-thesis] -->

@@

`r table_nums("datasets")`
```{r}
datastats %>%
  kable(format = "latex",
        booktabs = TRUE,
        col.names = c("Dataset", rep(c("Candidate studies", 
                                       #"Studies selected for fulltext screening", 
                                       "Relevant studies", 
                                       "Inclusion rate (%)"),2))) %>%
  kable_styling(full_width = TRUE) %>%
  add_header_above(c("", "Original study" = 3, "Test collection" = 3))
```

## Evaluating performance
Model performance was visualized by plotting recall curves and further assessed by three different measures, Work Saved over Sampling (WSS), Relevant References Found (RRF), and Average Time to Discovery (ATD). Results were averaged over 15 trials for every simulation.

Plotting recall as a function of the number of screened publications offers insight in model performance throughout the screening process [@Cormack2014; @Yu2018]. The curves give informations in two directions. On the one hand they display the number of publications that need to be screened to achieve a certain level of recall (WSS), but on the other hand they present how many relevant publications are identified after screening a certain proportion of all publications (RRF).

WSS indicates the reduction in publications needed to be screened, at a given level of recall [@Cohen2006]. 
When measured at the typical recall level of 0.95 [@Cohen2006], WSS yields an estimate of the amount of work that can have be saved at the cost of failing to identify 5% of relevant publications. In the current study, WSS is computed at 0.95 and 1.00 recall level. 
RRF statistics are computed at 10%, representing the proportion of relevant studies that were identified after screening 10% of all publications.

Both RRF and WSS are sensitive to random effects as these statistics are strongly dependent on the position of the cutoff value. 
Moreover, WSS makes assumptions about acceptable recall levels whereas this might depend on the research question at hand [@OMara-Eves2015]. 
A statistic that is not dependent on some arbitrary cutoff value is the ATD, which is the average number of publications that are screened to find a relevant publication, divided by the total number of publications in the data. The ATD is proportional to the area above the recall curve.

@@

### Optimizing hyperparameters
Every model component contains hyperparameters, leading to a unique set of hyperparameters for each model. To maximize model performance, we need to find optimal values for the hyperparameters. 
For every model the optimal hyperparameter values are determined by optimizing on the data $d$.
The hyperparameters are optimized by running several hundreds of optimization trials, in which hyperparameter values are sampled from their possible parameter space.
A description of all hyperparameters and their sample space can be found in the appendix.

>>>>

---
title: Manuscript drafts
author: "Gerbrich Ferdinands"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  word_document: default
  bookdown::pdf_document2:
    number_sections: no
    toc: no
bibliography: asreview.bib
link-citations: yes
csl: systematic-reviews.csl
---

<!-- --- -->
<!-- title: Manuscript drafts -->
<!-- author: "Gerbrich Ferdinands" -->
<!-- date: '`r format(Sys.time(), "%d %B, %Y")`' -->
<!-- output: -->

<!--   bookdown::pdf_document2: -->

<!--     number_sections: no -->
<!--     toc: no -->
<!--   word_document: default -->
<!-- bibliography: asreview.bib -->
<!-- link-citations: yes -->
<!-- csl: systematic-reviews.csl -->
<!-- --- -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(tinytex.verbose = TRUE)

library(tidyverse)
library(knitr)
library(xtable)
library(kableExtra)
library(captioner)
library(glue)
library(rbbt)
library(officer)
library(flextable)
# load model parameters 
source("asr-parameters.R")

@@

# Introduction
Systematic reviews are top of the bill in building evidence in research. A systematic review brings together all studies relevant to answer a specific research question [@PRISMA-PGroup2015]. Systematic reviews inform practice and policy [@Gough2002] and are key in developing clinical guidelines [@Chalmers2007]. 
<!-- PRISMA definition:  -->
<!-- > "A systematic review attempts to collate all relevant evidences that fits pre-specified eligibility criteria to answer a specific research -->
<!-- question. 
add that this is a systematic/rigorous process?-->

However, systematic reviews are costly because they involve the manual screening of thousands of titles and abstracts to identify publications relevant to answering the research question. An experienced reviewer takes on average 30 seconds to screen one title and abstract, whereas an inexperienced reviewer takes even longer [@Wallace2010]. Conducting a systematic review typically requires over a year of work by a team of researchers [@Borah2017]. 
<!-- Besides time-consuming, the process of title and abstract screening is prone to human error [@Wang2020]. -->
<!-- (14 to 77,910, trimmed mean 1,286).  -->
<!-- Need to be screened by multiple researchers.  --> 
Nevertheless, systematic reviewers are often bound to a limited budget and timeframe. Currently, the demand for systematic reviews exceeds the available time and resources by far [@Lau2019]. Especially when the need for guidelines is urgent - such as in the context of the current COVID-19 crisis - it is almost impossible to provide a review that is both timely and comprehensive. To ensure a timely review, reducing workload in systematic reviews is essential.

With advances in Artificial Intelligence (AI), there has been wide interest in tools to reduce workload in systematic reviews [@Harrison2020]. Various learning models have been proposed, aiming to predict whether a given publication is relevant or irrelevant to the systematic review. Findings suggest that such models potentially reduce workload with 30-70% at the cost of losing 5% of relevant publications, i.e. 95% recall [@OMara-Eves2015].
<!-- To date, several studies have begun to examine the use of learning algorithms for screening prioritization. -->

A well-established approach in increasing efficiency in title and abstract screening is screening prioritization [@Cohen2009; @Shemilt2014]. In screening prioritization, the learning model reorders publications to be screened by their likeliness to be relevant. The model presents the reviewer with the publications which are most likely to be relevant first, thereby expediting the process of finding all of the relevant publications. Such an approach allows for substantial time-savings in the screening process. Reviewing relevant publications early facilitates a faster transition of those publications to the next steps in the review process [@Cohen2009]. 
~~Additionally, although outside the scope of the current study, active learning models have the potential to reduce the number of publications needed to screen when combined with some sort of stopping criterion [@Yu2019].~~
<!-- Additionally, several studies report increasing efficiency beyond saving time [@OMara-Eves2015].  -->

<!-- assist the reviewer in the screening process -->
<!-- Several studies have proposed active learning models for screening prioritization -->
Recent studies have demonstrated the effectiveness of screening prioritization by means of active learning models [@Yu2018; @Yu2019; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010]. Active learning is when the model can iteratively improve its predictions by allowing the model to choose the data from which it can learn [@Settles2012]. Active learning has proven to be an efficient strategy in large datasets where labels are scarce, which makes title and abstract screening an ideal candidate for such models. When applied in screening prioritization, the reviewer screens publications that are presented by an active learning model. Subsequently, the active learning model learns from the reviewers' decision ('relevant', 'irrelevant') and uses this knowledge in selecting the next publication to be screened by the reviewer.

<!-- Combined with some sort of stopping criterion, the reviewer c -->
<!-- _Adopting some sort of stopping criterion (outside scope of the current thesis) the reviewer can quit reviewing after having read only a fraction of candidate studies. meaning the screening process can be finished after reading a fraction of all candidate studies. Saving hours of time and resources._ -->

Although the application of learning models in reducing workload of systematic reviews has been extensively studied [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010], the complex nature of the field is making it difficult to draw overarching conclusions about best practice [@OMara-Eves2015]. First, whilst previous studies have evaluated models in many forms and shapes [@Yu2018; @Yu2018; @Miwa2014; @Wallace2010], all models adopt the same classification technique. Even though a plethora of classification techniques exists [@Aggarwal2012], no study in this area has made a comparison between models adopting different classification techniques. 
<!-- whilst there exists a  of classification techniques for an active learning model to adopt, research to date has confined to models adopting one classification strategy [@Yu2018; @Yu2018; @Miwa2014; @Wallace2010]. Models adopting other classification techniques have not been investigated.  -->
 <!-- evaluated techniques other than Support Vector Machines (SVM) -->
Second, the lack of model replication on reviews from varying research contexts makes it impossible to draw conclusions about the general effectiveness of active learning models [@OMara-Eves2015; @Marshall2020]. As far as known to the authors, Miwa et al [@Miwa2014] were the only researchers to make a direct comparison between systematic reviews from different research areas, namely the social and the medical sciences. They found that active learning was more difficult on data from the social sciences due to the nature of the different vocabularies used. Therefore, it is of interest to evaluate model performance across different research contexts.

<!-- Taken together, the question remains how active learning models for screening prioritization perform across different classification techniques (1) and review contexts (2). Hence, additional evaluations of active learning models are required. The current study aims to address these issues by answering the following research questions:  -->
Taken together, evaluations of active learning models are required across different classification techniques (1) and review contexts (2).  The current study aims to address these issues by answering the following research questions:

  __RQ1__ What is the performance of active learning models across different classification techniques?

  __RQ2__ Does performance of active learning models differ across systematic reviews from different research areas?

<!-- Although important issues in the application of active learning models on title and abstract screening have been addressed prior studies -->

@@

The purpose of the current paper is to increase the evidence base on active learning models for reducing workload in title and abstract screening in systematic reviews. Combining latest insights from this area, we propose seven different active learning models for the purpose of identifying relevant publications in systematic review datasets. The models were chosen to maximize the number of identified relevant publications, while minimizing the number of publications needed to screen. 
Working towards a general consensus in this emerging field, model performance was assessed by conducting a retrospective simulation on six systematic review datasets. Datasets were collected from the fields of medicine, software engineering, psychology, behavioural public administration, and virology to assess generalizability of the models across research contexts. The models, datasets and simulations are implemented in a pipeline of active learning for screening prioritization, called $\texttt{ASReview}$ [@ASReview2020]. $\texttt{ASReview}$ is an open source and generic tool such that users can adapt and add modules as they like, encouraging fellow researchers to replicate findings from previous studies. All scripts and data used are openly published to facilitate usability and acceptability of AI-assisted title and abstract screening in the field of systematic review.

\textcolor{red}{The remaining part of this paper is organized as follows. The Methods section elaborates on the setup of the current study, starting with a task description, followed the Technical Details section which will cover the workings of active learning models for study selection in systematic reviews on a conceptual level.
The results section reports ... The discusssion section summarises the findings, comments on them and  summarises the main findings, discusses 
discuss limitations, draw conclusion and}

@@

# Methods 
## Task Description
The screening process of a systematic review starts with all publications obtained in the search. The task is to identify which of these publications are relevant, by screening them at the title and abstract level. In active learning for screening prioritization, the screening process proceeds as follows:

- Start with the set of all unlabeled (titles and) abstracts, $\mathcal{U}$.
- The reviewer provides a label for a few records $x \in \mathcal{U}$, creating an set of labeled titles and abstracts $\mathcal{L}$. The label can be either relevant $x_R$ or irrelevant $x_I$.
- The active learning cycle starts:

  1. A classifier is trained on the labeled titles and abstracts, $C = \texttt{train}(\mathcal{L})$

  2. The classifier predicts labels for all unlabelled titles and abstracts, $C(\mathcal{U})$

  3. Based on the predictions by $C$, the model selects the most relevant title and abstract $x^* \in \mathcal{U}$

  4. The model asks the reviewer to screen this title and abstract, $x^*_?$

  5. The reviewer screens the title and abstract and provides a label, $x^*_R$ or $x^*_I$

  6. The labeled title and abstract is added to the training data, $x_{R \texttt{ xor } I} \in \mathcal{L}$

@@

- In this active learning cycle, the model can incrementally improve its predictions on the remaining unlabeled title and abstracts. The relevant titles and abstracts are identified as early in the process as possible. The reviewer and the model keep interacting until the reviewer decides to stop or until all title and abstracts have been labelled.

@@

### Classification
To make predictions on the unlabeled publications, a classifier is trained on features from the set of previously labeled publications. 
<!-- The classifier predicts class of a given unlabeled publication.  -->
A technique widely used in classification tasks is the Support Vector Machine (SVM). SVMs separate the data into classes by finding a multidimensional hyperplane [@Tong2001; @Kremer2014]. SVMs have been proven to be effective in active learning models for screening prioritization [@Yu2018; Miwa2014]. Moreover, SVMs are the currently the only classifier implemented in ready-to-use software tools implementing active learning for screening prioritization (Abstrackr [@Wallace2012b], Colandr [@Cheng2018], FASTREAD [@Yu2018], Rayyan [@Ouzzani2016], and RobotAnalyst [@Przybyla2018]).

Whilst the performance of several classification techniques has been investigated in the AI-aided title-and-abstract screening field in general [@Kilicoglu2009; @Aphinyanaphongs2004], the relatively new subfield of active learning for screening prioritization has not yet studied the performance of classifiers other than SVMs [@Yu2018; @Yu2018; @Cormack2014; @Cormack2015; @Miwa2014; @Wallace2010]. 
<!-- Whilst the SVM is a well-known technique, also classifiers have been employed in class prediction tasks in systematic reviews.  -->
<!-- In fact, research in this area to date has not yet determined the performance of classifiers other than SVMs. -->
The current study aims to address this gap by exploring performance of three classifiers besides SVM:

@@

These three classification techniques were selected because they are widely adopted methods in text classification [@Aggarwal2012]. Moreover, these techniques can be run on a personal computer as they require a relatively low amount of processing power.

@@

### Class imbalance problem
There are two classes in the dataset: relevant and irrelevant publications. Typically, the inclusion rate is low (<10%) as only a fraction of the publications belong to the relevant class. This poses a problem for training a classifier as there are far fewer examples of relevant than irrelevant publications to train on [@OMara-Eves2015]. Moreover, classifiers are well-suited to separate data into classes, but not to correctly identify one class [@Wallace2010]. Therefore, the class imbalance problem causes the classifier to miss relevant publications. This is evident in the case of a systematic review dataset where only one percent of publications are relevant. A model would achieve 99% accuracy when classifying all publications as irrelevant, even though none of the relevant papers would have been correctly identified.

Previous studies have addressed the class imbalance problem by rebalancing the training data in different ways [@OMara-Eves2015]. To decrease the class imbalance in the training data, the models in the current study rebalance the training set by Dynamic Supersampling (DS). DS decreases the number of irrelevant publications in the training data, whereas the number of relevant publications are increased (by copy) such that the size of the training data remains the same. The ratio between relevant and irrelevant publications in the training data is not fixed, but dynamically updated and depends on the size of the training data, the total number of publications, and the ratio between the total number of labeled publications.

### Word embeddings 
To predict publication class, the classifier uses information from the publications in the dataset. 
Examples of such information are titles and abstracts. However, a model cannot make predictions from the titles and abstracts as they are; their textual content needs to be represented numerically. The textual information needs to be mapped to feature vectors. This process of numerically representing textual content is called 'word embeddings'.

A classical example of word embeddings is a 'bag of words' (bow) representation. For each text in the data set, the number of occurrences of each word is stored. This leads to $n$ features, where $n$ is the number of distinct words in the texts [@scikit-learn]. The bag-of-words method is simplistic and will highly value often occurring but otherwise meaningless words such as "and". A more sophisticated approach is Term-Frequency Inverse Document Frequency (TF-IDF). TF-IDF circumvents this problem by adjusting the term frequency in a text with the inverse document frequency, the frequency of a given word in the entire data set [@Ramos2003]. A downside of TF-IDF and other bow methods is that they do not take into account the ordering of words, thereby ignoring semantics. An example of an approach that aims to overcome this weakness is Doc2Vec (D2V), capable of grasping the relations between words by learning to predict the words in the texts [@Le2014].

Miwa et al. found that active learning was more difficult on data from the social sciences compared to data from the medical sciences and were able to link this difficulty to a natural difference in text complexity between these research areas [@Miwa2014]. As the study by Miwa et al. adopted a bow approach [@Miwa2014], we hypothesize that a more sophisticated word embedding strategy has the potential to bridge the performance gap between these research areas.

### Query strategy
The active learning model can adopt different strategies in selecting the next publication to be screened by the reviewer. 
A strategy mentioned before is selecting the publication with the highest probability of being relevant. In the active learning literature this is referred to as certainty-based active learning [@Settles2012]. Another well-known strategy is uncertainty-based active learning, where the instances that will be presented next will be those instances on which the model's classifications are the least certain, i.e. close to 0.5 probability [@Settles2012]. Traditionally, this strategy trains the most accurate model because the model can learn the most from instances it is uncertain about. However, a study comparing performance of both strategies in detecting relevant publications found that the accuracy gain of uncertainty-based screening was not significant [@Miwa2014].

Certainty-based active learning is the preferred strategy for the task at hand. Firstly, this strategy is far better suited to the goal of prioritizing relevant publications compared to uncertainty-based active learning, in which the publications are prioritized that the model is most uncertain about. Secondly, certainty-based active learning is far better equipped at dealing with imbalanced data in active learning [@Fu2011].

<!-- most equipped to the current scenario in which we are dealing with highly imbalanced datasets and where the goal is to identify all relevant publications as soon as possible. -->

@@

First, four models combining every classifier with TF-IDF word embeddings were investigated:

 - SVM + TF-IDF
 - NB + TF-IDF
 - RF + TF-IDF 
 - SVM + TF-IDF

Second, the classifiers were combined with Doc2Vec word embeddings, leading to the following three models:^[The combination NB + D2V could not be tested because the Multinomial Naive Bayes classifier (sklearn) can only handle a feature matrix with positive values, whereas the doc2vec word embeddings approach (gensim) produces a feature matrix that can also contain negative values.]

@@

## Simulation study
For each of the seven models, performance was evaluated by simulating the model on the screening process of six systematic reviews. 
Performance of the seven models was evaluated by simulating their behaviour in the screening process of six systematic review datasets.
Put differently, 42 simulations were carried out. For every model-dataset combination, hyperparameters were optimized^[see the Appendix for more information]. To account for variance, every simulation was repeated for 15 trials. 
Simulations were run using \texttt{ASReview}'s simulation mode [@ASReview2020]. There was no need for a human reviewer as the model could query the labels in the data instead.

Every simulation started with an initial training set of one relevant and one irrelevant publication to represent a 'worst case scenario' where the reviewer has minimal prior knowledge on the publications in the data. To account for bias, the initial training set was randomly sampled from the dataset for every of the 15 runs. Although varying over runs, the initial training sets were kept constant over datasets to allow for a direct comparison of models within datasets. A seed value was set to ensure reproducibility 
The classifier was retrained every time after a publication had been labeled. 
The simulation ended after all publications in the dataset had been labeled.

@@

## Datasets 
The models were simulated on a convenience sample of six systematic review datasets. The data selection process was driven by two factors. Firstly, datasets were selected based on their background, given the need for datasets from diverse research areas. Secondly, datasets were selected by their availability, given the limited timespan of the current project. The datasets were retrieved from a collection of open systematic review datasets to be used for text mining purposes^[https://github.com/asreview/systematic-review-datasets].

@@

Data were preprocessed from their original source into a test dataset, containing title and abstract of the publications obtained in the initial search. Candidate studies with missing abstracts and duplicate instances were removed from the data. Test datasets were labelled to indicate which candidate studies were included in the systematic review, thereby indicating relevant publications. All test datasets consisted of thousands of candidate studies, of which only only a fraction was deemed relevant to the systematic review. For four of the six datasets inclusion rates were centered around 1-2 percent. For the remaining two datasets, the inclusion rate was about 5 percent (`r table_nums("datasets", display = "cite")`). 
<!-- The data preprocessing scripts can be found on the GitHub^[https://github.com/GerbrichFerdinands/asreview-thesis] -->

@@

`r table_nums("datasets")`
```{r}
datastats %>%
  kable(format = "pandoc",
        booktabs = TRUE,
        col.names = c("Dataset", rep(c("Candidate studies", 
                                       #"Studies selected for fulltext screening", 
                                       "Relevant studies", 
                                       "Inclusion rate (%)"),2))) %>%
  kable_styling(full_width = TRUE) %>%
  add_header_above(c("", "Original study" = 3, "Test collection" = 3))
```

## Evaluating performance
Model performance assessed by three different measures, Work Saved over Sampling (WSS), Relevant References Found (RRF), and Average Time to Discovery (ATD). Furthermore, model performance was visualized by plotting recall curves. Results were averaged over 15 trials for every simulation.

WSS indicates the reduction in publications needed to be screened, at a given level of recall [@Cohen2006]. Typically measured at a recall level of 0.95 [@Cohen2006], WSS@95 yields an estimate of the amount of work that can be saved at the cost of failing to identify 5% of relevant publications. In the current study, WSS is computed at 0.95 and 1.00 recall level. 
RRF statistics are computed at 10%, representing the proportion of relevant studies that are after screening 10% of all publications.

Both RRF and WSS are sensitive to random effects as these statistics are strongly dependent on the position of the cutoff value. Moreover, WSS makes assumptions about acceptable recall levels whereas this level might depend on the research question at hand [@OMara-Eves2015].

A statistic that is not dependent on some arbitrary cutoff value is the TTD, which is the average number of publications that needed to be screened to find a relevant publication, divided by the total number of publications in the data. The TTD is proportional to the area above the recall curve.

Plotting recall as a function of the number of screened publications offers insight in model performance throughout the screening process [@Cormack2014; @Yu2018]. The curves give information in two directions. On the one hand they display the number of publications that need to be screened to achieve a certain level of recall (WSS), but on the other hand they present how many relevant publications are identified after screening a certain proportion of all publications (RRF).

@@

<!-- ### Optimizing hyperparameters -->
<!-- Every model component contains hyperparameters, leading to a unique set of hyperparameters for each model. To maximize model performance, we need to find optimal values for the hyperparameters.  -->
<!-- For every model the optimal hyperparameter values are determined by optimizing on the data $d$. -->
<!-- The hyperparameters are optimized by running several hundreds of optimization trials, in which hyperparameter values are sampled from their possible parameter space. -->
<!-- A description of all hyperparameters and their sample space can be found in the appendix.  -->
